RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-08-15 jsac.tex
:triangular_flag_on_post: JSAC 230815

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
% \usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{setspace}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}

% page limit hacks
% ! \usepackage[top=1cm, bottom=1cm, left=1cm, right=1cm]{geometry}
% \abovedisplayskip=1mm
% \belowdisplayskip=1mm
% \abovedisplayshortskip=1mm
% \belowdisplayshortskip=1mm
\setlength{\jot}{0.1mm}
\setlength{\floatsep}{1mm}
\setlength{\textfloatsep}{1mm}
\setlength{\intextsep}{1mm}
\setlength{\skip\footins}{2mm}
% dark mode
% \usepackage{xcolor} \pagecolor[rgb]{0,0,0} \color[rgb]{1,1,1}

@@

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communication}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communication}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmmac}{DMMAC}{Discrete Memoryless Multiple Access Channel}
\newacronym{dcmc}{DCMC}{Discrete-input Continuous-output Memoryless Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic \glsentryshort{bc}}
\newacronym{bbc}{BBC}{Bistatic \glsentryshort{bc}}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}
\newacronym{ber}{BER}{Bit Error Rate}

\begin{document}
% \title{RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface From a Distribution Perspective}
% \title{RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface}
% \title{RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface From State Distribution}
% \title{RIScatter: Bridging Backscatter Modulation and Passive Beamforming With Adaptive State Distribution}
\title{RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	\thanks{
		The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
		B. Clerckx is also with Silicon Austria Labs (SAL), Graz A-8010, Austria.
	}
}
\maketitle
% TODO why not use metamaterial? because high computational complexity and expectations of freedom (flexibility, reliability)
% * cooperative input design can be used for co-located RIScatter nodes.
% * hardward to modulate using RIS is already available.
% * RIScatter: enhance backscatter signal characteristics (suppress primary interference) or primary signal strength.
% ? For Backscatter community: why not adaptive channel coding? For RIS community: why not add randomness for modulation?
% ? RIS involves nonlinear circuits (e.g., diodes) and impedance design (i.e., perfect mismatching) is tricky for continuous phase shift; discrete candidate set is preferred (2-3 bits optimal)
% * Redefine RIS using book
% x Using RIS to encode is not very interesting as controller is anyway usable
% * Wi-Fi sampling rate 20 MHz (symbol period 50 nS), RFID symbol rate/load-switching speed 100s KHz to 10s MHz [Backscatter Communication], fast load-switching speed is key for high data rate - symbol ratio is not very large
% * Beamforming can cancel direct interference to avoid the collapse of noncoherent energy detection and enlarge energy gap.
% ? [Reconfigurable Intelligent Surface-Assisted Ambient Backscatter Communication Experimental Assessment]the authors of [5] show that, even in line-of-sight (LOS), the interferences and the desired signal can combine in such a way that the received signals at the reader side in the two states are close in power and only differ by their phases
% TODO Input probability distribution for discrete-value reflection (backscatter), constellation design for continue-value reflection (RIS)
% * channel coding: non-adaptive (line-coding) or adaptive (DMC)
% ? [Turbocharging] Since averaging must be done over a significant period of time to smooth out the variation in the TV transmissions, averaging does not allow rates much higher than 10 kbps [24].
% ? [mu-code]: use multiple Rxs for noncoherent direct interference cancellation (instead of receive diversity)
% * backscatter modulation <-> passive beamforming; add adaptive channel coding over (NRZ) line coding.
% ? semi-coherent: known CSI, unknown primary/backscatter symbols
% * initialization, convergence, base 2 in simulation
% * SR analysis can achieve large rate at the cost of large N; RIScatter can achieve slightly lower rate at the same N, but also preserves primary performance for smaller N
% TODO tradeoff between primary (RIScatter, no SIC) and backscatter (SR, SIC)
% * SR assume very large N for primary rate analysis. However, in such cases SIC is questionable as energy detection over received signal as BBC is already good enough and noncoherent detection can also achieve reasonable performance?
% x when primary symbols are determined, optimal input distribution at high backscatter SNR is equiprobable for any detection (signal/energy)
% ? Channel estimation error, coordination issues

% * add table row: adaptive to information source/csi/link weights
% * explain link weights
% * add table row: distributions
% * table to sentences
% * add primary and backscatter flows to illustrations
% * discuss table and add challenges in contribution 1
% * transition challenges (distribution and constellation-adaptive) to contribution 2, 3
% * think RIScatter applications (use Rui's slides)

\begin{abstract}
	% \gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.
	% \gls{sr} incorporates a passive scatter node into existing radio networks to create additional propagation path and ride its own information towards the cooperative receiver.
	% \gls{ris} is a programmable reflector array that adapts the phase shift response to enhance or suppress signal strength in specific directions.
	\gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave.
	\gls{ris} adapts its phase shift response to enhance or attenuate channel strength in specific directions.
	In this paper, we show how those two seemingly different technologies (and their derivatives) can be unified into a single architecture called RIScatter.
	RIScatter consists of multiple dispersed or co-located scatter nodes, whose reflection states are adapted to partially modulate their own information and partially engineer the wireless channel.
	% This contrasts with \gls{bc} (resp. \gls{ris}) where the reflection pattern is exclusively a function of the information symbol (resp. \gls{csi}).
	The key principle is to render the probability distribution of reflection states as a joint function of the information source, \gls{csi}, and \gls{qos} of coexisting active primary and passive backscatter links.
	This enables RIScatter to softly bridge \gls{bc} and \gls{ris}; boil down to either under specific input distribution; or evolve in a mixed form for heterogeneous traffic control and universal hardware design.
	% To accommodate the characteristics of active-passive coexisting networks, we propose a practical \gls{sic}-free receiver
	To reap the benefits of RIScatter, we also propose a co-located \gls{sic}-free receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	For a single-user multi-node RIScatter network, we characterize the achievable primary-(total-)backscatter rate region by designing the input distribution at scatter nodes, the active beamforming at the \gls{ap}, and the energy decision regions at the user.
	Simulation results demonstrate RIScatter nodes can exploit the scattered paths to smoothly shift between backscatter modulation and passive beamforming.
\end{abstract}

\begin{IEEEkeywords}
	Input distribution design, symbol-level precoding, index modulation, active-passive coexisting network, \gls{sic}-free receiver, energy detection, backscatter communication, reconfigurable intelligent surface, ambient backscatter communication, symbiotic radio.
\end{IEEEkeywords}

@@

\begin{section}{Introduction}
	% \begin{subsection}{Fundamentals}
	\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of low-power devices.
	% As a mature low-power communication technique,
	\gls{bc} separates a transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
	% In particular, the node harvests energy from impinging wave and embeds information over scattered signal, and
	The receiver (reader) can be either co-located or separated with the carrier emitter, known as \gls{mbc} and \gls{bbc} in Fig.~\subref*{fg:mbc} and \subref*{fg:bbc}, respectively.
	Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized to embrace the \gls{ioe}.
	However, conventional backscatter nodes only respond when externally inquired by a nearby reader.
	% To tackle this, \cite{Liu2013b} proposed \gls{ambc} where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connection in between.
	\gls{ambc} in Fig.~\subref*{fg:ambc} was proposed a decade ago where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connections \cite{Liu2013b}.
	It does not require dedicated power source, carrier emitter, or frequency spectrum, but the backscatter decoding is subject to the strong interference from the primary (legacy) link.
	% Cooperative \gls{ambc} that employs a co-located receiver to decode both coexisting links was considered in \cite{Yang2018}, and the rate and error performance are evaluated for various detection schemes.
	To tackle this, cooperative \gls{ambc} \cite{Yang2018} employs a co-located receiver to decode both coexisting links and the concept was further refined as \gls{sr} in Fig.~\subref*{fg:sr} \cite{Liang2020}.
	% Specifically, the active transmitter generates \gls{rf} wave carrying primary information, the passive node ride extra information over the scattered signal, and the cooperative receiver jointly, individually, or sequentially decodes both links.
	Specifically, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates a rich-scattering environment and rides its own information, and the co-located receiver cooperatively decodes both links.
	In those \gls{bc} applications, the scatter node is considered as an \emph{information source} and the reflection pattern depends exclusively on the information symbol.
	% Those \gls{bc} applications employ scatter nodes as pure information sources, and the active primary transmission (if it exists) can be influenced by the randomness of backscatter modulation.
	On the other hand, \gls{ris} in Fig.~\subref*{fg:ris} is a smart signal reflector with numerous passive elements of adjustable phase shifts.
	It customizes the wireless environment for signal enhancement, interference suppression, scattering enrichment, and/or non-line-of-sight bypassing \cite{Wu2021b}.
	Each \gls{ris} element is considered as a \emph{channel adaptor} and the reflection pattern depends exclusively on the \gls{csi}.
	% In summary, \gls{mbc}/\gls{bbc} (resp. \gls{ris}) involves a \emph{single} passive backscatter (resp. active primary) link where the passive device modulates information (resp. enhances propagation) and the active device generates carrier wave (resp. modulated signal).
	% On the other hand, \gls{ambc} (resp. \gls{sr}) involves \emph{coexisting} active primary and passive backscatter links in a competitive (resp. collaborative) relationship.
	% In \gls{ambc}, the backscatter link only needs to know the basic information (e.g., operating frequency) of the primary link, while the primary link may be completely unaware of the backscatter link.
	% In \gls{sr}, relevant \gls{csi} is exploited at the co-located receiver for cooperative decoding, and may also be exploited at the primary transmitter for cooperative active beamforming.

	\begin{figure}[!t]
		\centering
		\subfloat[\gls{mbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/mbc.tex}
			}
			\label{fg:mbc}
		}
		\subfloat[\gls{bbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/bbc.tex}
			}
			\label{fg:bbc}
		}
		\subfloat[\gls{ambc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ambc.tex}
			}
			\label{fg:ambc}
		}
		\\
		\subfloat[\gls{sr}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/sr.tex}
			}
			\label{fg:sr}
		}
		\subfloat[\gls{ris}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ris.tex}
			}
			\label{fg:ris}
		}
		\subfloat[RIScatter]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/riscatter.tex}
			}
			\label{fg:riscatter}
		}
		\caption{
			Illustration of scattering applications.
			The blue flow(s) constitutes the primary link while the magenta/green flow denotes the backscatter link.
		}
		\label{fg:scatter_illustration}
	\end{figure}
	% \end{subsection}

	\begin{table*}[!t]
		\caption{Comparison of Scattering Applications}
		\label{tb:scattering_applications}
		\rowcolors{2}{gray!25}{white}
		\scriptsize
		\begin{tabularx}{\textwidth}{CCCCCC}
			\toprule
			\hiderowcolors
			                                       & \gls{mbc}/\gls{bbc} & \gls{ambc}                  & \gls{sr} (large spreading factor)               & \gls{ris}           & RIScatter                                                   \\ \midrule
			\showrowcolors
			Information link(s)                    & Backscatter         & Coexisting                  & Coexisting                                      & Primary             & Coexisting                                                  \\
			Primary signal on backscatter decoding & Carrier             & Multiplicative interference & Spreading code                                  & ---                 & Energy uncertainty                                          \\
			Backscatter signal on primary decoding & ---                 & Multiplicative interference & \gls{csi} uncertainty                           & Passive beamforming & Dynamic passive beamforming                                 \\
			Cooperative devices                    & ---                 & No                          & Primary transmitter and co-located receiver     & ---                 & Primary transmitter, scatter nodes, and co-located receiver \\
			Sequential decoding                    & ---                 & No                          & Primary-to-backscatter, \gls{sic} and \gls{mrc} & ---                 & Backscatter-to-primary, no \gls{sic}/\gls{mrc}              \\
			% Primary decoding                        & ---                 & Semi-coherent               & Semi-coherent                                   & Coherent            & Coherent                                                    \\
			% Backscatter decoding                    & Coherent            & Semi-coherent               & Coherent                                        & ---                 & Semi-coherent                                               \\
			Reflection pattern depends on          & Information source  & Information source          & Information source                              & \gls{csi}           & Information source, \gls{csi}, and \gls{qos}                \\
			Reflection state distribution          & Equiprobable        & Equiprobable                & Equiprobable or Gaussian                        & Degenerate          & Flexible                                                    \\
			Load-switching speed                   & Fast                & Slow                        & Slow                                            & Quasi-static        & Arbitrary                                                   \\ \bottomrule
		\end{tabularx}
	\end{table*}

>>>>

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[makeroom]{cancel}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
% \usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}

% page limit hacks
\usepackage{setspace}
% ! \usepackage[top=1cm, bottom=1cm, left=1cm, right=1cm]{geometry}
% \abovedisplayskip=1mm
% \belowdisplayskip=1mm
% \abovedisplayshortskip=1mm
% \belowdisplayshortskip=1mm
\setlength{\jot}{0.1mm}
\setlength{\floatsep}{1mm}
\setlength{\textfloatsep}{1mm}
\setlength{\intextsep}{1mm}
\setlength{\skip\footins}{2mm}

@@

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communication}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communication}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{clt}{CLT}{Central Limit Theorem}
\newacronym{cw}{CW}{Continuous Waveform}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{dc}{DC}{Direct Current}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmmac}{DMMAC}{Discrete Memoryless Multiple Access Channel}
\newacronym{dcmc}{DCMC}{Discrete-input Continuous-output Memoryless Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pga}{PGA}{Projected Gradient Ascent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic \glsentryshort{bc}}
\newacronym{bbc}{BBC}{Bistatic \glsentryshort{bc}}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}
\newacronym{ber}{BER}{Bit Error Rate}

\begin{document}
\title{RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	\thanks{
		The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
		B. Clerckx is also with Silicon Austria Labs (SAL), Graz A-8010, Austria.
	}
}
\maketitle

\begin{abstract}
	\gls{bc} nodes harvest energy from and modulate information over external electromagnetic waves.
	\gls{ris} adapts its phase shift response to alter channel strength in specific directions.
	In this paper, we show how those two seemingly different technologies (and their derivatives) can be unified into one architecture called RIScatter.
	RIScatter consists of dispersed or co-located scatter nodes, whose reflection states are adapted to partially modulate their information and partially engineer the wireless channel.
	The key is to render the probability distribution of reflection states as a joint function of the information source, \gls{csi}, and relative priority of coexisting links.
	This enables RIScatter to softly bridge \gls{bc} and \gls{ris}; reduce to either under specific setup; or evolve in a mixed form for heterogeneous traffic control and universal hardware design.
	We also propose a low-complexity \gls{sic}-free receiver that exploits the properties of RIScatter.
	For a single-user multi-node network, we characterize the achievable primary-(total-)backscatter rate region by optimizing the input distribution at scatter nodes, the active beamforming at the \gls{ap}, and the energy decision regions at the user.
	Simulations demonstrate RIScatter nodes can recycle surrounding radios for backscatter modulation and passive beamforming.
\end{abstract}

\begin{IEEEkeywords}
	Backscatter communication, reconfigurable intelligent surface, active-passive coexisting network, input distribution design, \gls{sic}-free receiver.
\end{IEEEkeywords}

@@

\begin{section}{Introduction}
	\label{sc:introduction}
	\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of low-power devices.
	\gls{bc} separates a transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
	The receiver (reader) can be either co-located or separated with the carrier emitter, known as \gls{mbc} and \gls{bbc} in Fig.~\subref*{fg:mbc} and \subref*{fg:bbc}, respectively.
	Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized to embrace the \gls{ioe}.
	However, conventional backscatter nodes only respond when externally inquired by a nearby reader.
	\gls{ambc} in Fig.~\subref*{fg:ambc} was proposed a decade ago where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connections \cite{Liu2013b}.
	It does not require dedicated power source, carrier emitter, or frequency spectrum, but the backscatter decoding is subject to the strong interference from the primary (legacy) link.
	To tackle this, cooperative \gls{ambc} \cite{Yang2018} employs a co-located receiver to decode both coexisting links, and the concept was further refined as \gls{sr} in Fig.~\subref*{fg:sr} \cite{Liang2020}.
	Specifically, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates a rich-scattering environment and superimposes its own information, and the co-located receiver cooperatively decodes both links.
	In those \gls{bc} applications, the scatter node is considered as an \emph{information source} and the reflection pattern depends exclusively on the information symbol.
	On the other hand, \gls{ris} in Fig.~\subref*{fg:ris} is a smart signal reflector with numerous passive elements of adjustable phase shifts.
	It customizes the wireless environment for signal enhancement, interference suppression, scattering enrichment, and/or non-line-of-sight bypassing \cite{Wu2021b}.
	Each \gls{ris} element is considered as a \emph{channel shaper} and the reflection pattern depends exclusively on the \gls{csi}.

	As a special case of \gls{cr}, active and passive transmissions coexist and interplay in \gls{ambc} and \gls{sr}.
	Such a coexistence is classified into commensal (overlay), parasitic (underlay), and competitive (interfering) paradigms, and their achievable rate and outage performance were investigated in \cite{Guo2019b,Ding2020}.
	The achievable rate and optimal input distribution for binary-input \gls{ambc} are investigated in \cite{Qian2019b}, but its impact on the primary link is omitted.
	% For binary-input \gls{ambc},  are investigated in \cite{Qian2019b}, but the authors did not consider its impact on the primary link.
	% The \gls{ber} performance of \gls{ml}, linear, and \gls{sic} receivers are derived over flat fading channels \cite{Yang2018}.
	% The authors of \cite{Hassani2023} analyzed the energy efficiency and achievable rate region for an \gls{ambc}-aided multi-user downlink \gls{noma} system.
	In \cite{Hassani2023}, the authors analyzed the energy efficiency and achievable rate region for an \gls{ambc}-aided multi-user downlink \gls{noma} system.
	However, they assumed equal symbol duration and perfect synchronization for the coexisting links.
	Importantly, active-passive coexisting networks have three special and important properties:
	\begin{enumerate}
		\item Primary and backscatter symbols are superimposed by \emph{double modulation} (i.e., multiplication coding);
		\item Backscatter signal strength is much weaker than primary due to \emph{double fading};
		\item The spreading factor (i.e., backscatter symbol duration over primary) is usually large\footnote{The load-switching interval of low-power backscatter modulators is usually \num{0.1} to \qty{10}{\us} \cite{Torres2021}, accounting for a typical spreading factor between \num{10} and \num{e3}.}.
	\end{enumerate}
	The second property motivated \cite{Long2020a,Liang2020,Guo2019b,Ding2020,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Yang2018,Han2021,Zhang2022,Dai2023} to view \gls{sr} as a multiplicative \gls{noma} and perform \gls{sic} from primary to backscatter link.
	During primary decoding, the backscatter signal can be modelled as channel uncertainty or multiplicative interference, when the spreading factor is large or small.
	Decoding each backscatter symbol also requires multiple \gls{sic} followed by a \gls{mrc} over primary blocks, which is operation-intensive and \gls{csi}-sensitive.
	Under those assumptions, the achievable rate region of cell-free \gls{sr} is characterized in \cite{Dai2023}.
	When the spreading factor is sufficiently large, the primary achievable rate under semi-coherent detection\footnote{In this paper, semi-coherent detection refers to the primary/backscatter decoding with known \gls{csi} and unknown backscatter/primary symbols.} asymptotically approaches its coherent counterpart such that both links are approximately interference-free \cite{Long2020a}.
	However, this assumption severely limits the backscatter throughput.

	\begin{figure}[!t]
		\centering
		\subfloat[\gls{mbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/mbc.tex}
			}
			\label{fg:mbc}
		}
		\subfloat[\gls{bbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/bbc.tex}
			}
			\label{fg:bbc}
		}
		\subfloat[\gls{ambc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ambc.tex}
			}
			\label{fg:ambc}
		}
		\\
		\subfloat[\gls{sr}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/sr.tex}
			}
			\label{fg:sr}
		}
		\subfloat[\gls{ris}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ris.tex}
			}
			\label{fg:ris}
		}
		\subfloat[RIScatter]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/riscatter.tex}
			}
			\label{fg:riscatter}
		}
		\caption{
			Illustration of scattering applications.
			The blue flow(s) constitutes the primary link while the magenta/green flow denotes the backscatter link.
		}
		\label{fg:scatter_illustration}
	\end{figure}

	\begin{table*}[!t]
		\caption{Comparison of Scattering Applications}
		\label{tb:scattering_applications}
		\rowcolors{2}{gray!25}{white}
		\scriptsize
		\begin{tabularx}{\textwidth}{CCCCCC}
			\toprule
			\hiderowcolors
			                                       & \glsentryshort{mbc}/\glsentryshort{bbc} & \glsentryshort{ambc}        & \glsentryshort{sr} (large spreading factor)                         & \glsentryshort{ris} & RIScatter                                                          \\ \midrule
			\showrowcolors
			Information link(s)                    & Backscatter                             & Coexisting                  & Coexisting                                                          & Primary             & Coexisting                                                         \\
			Primary signal on backscatter decoding & Carrier                                 & Multiplicative interference & Spreading code                                                      & ---                 & Energy uncertainty                                                 \\
			Backscatter signal on primary decoding & ---                                     & Multiplicative interference & \glsentryshort{csi} uncertainty                                     & Passive beamforming & Dynamic passive beamforming                                        \\
			Cooperative devices                    & ---                                     & No                          & Primary transmitter and co-located receiver                         & ---                 & Primary transmitter, scatter nodes, and co-located receiver        \\
			Sequential decoding                    & ---                                     & No                          & Primary-to-backscatter, \glsentryshort{sic} and \glsentryshort{mrc} & ---                 & Backscatter-to-primary, no \glsentryshort{sic}/\glsentryshort{mrc} \\
			Reflection pattern depends on          & Information source                      & Information source          & Information source                                                  & \glsentryshort{csi} & Information source, \glsentryshort{csi}, and \glsentryshort{qos}   \\
			Reflection state distribution          & Equiprobable                            & Equiprobable                & Equiprobable or Gaussian                                            & Degenerate          & Flexible                                                           \\
			Load-switching speed                   & Fast                                    & Slow                        & Slow                                                                & Quasi-static        & Arbitrary                                                          \\ \bottomrule
		\end{tabularx}
	\end{table*}
